---
title: "Class 5"
author: "Emily Arnold"
date: "2/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
#have to install plotly
library(plotly)
```

Plotly!
-Interactive visualizations without actual coding
-Written in javascript

Working with the diamonds dataset:
-hover over to see individual points
-zoom in

```{r checking out dataset - Diamonds}
head(diamonds)
diamonds.sample <- diamonds %>% 
  sample_n(10000)

```

Couple different ways to interface with plotly:

```{r}

plot_ly(data = diamonds, x = ~carat, y = ~price)
plot_ly(data = diamonds, x = ~cut, y = ~price)

```

You could also do this tidyverse style

```{r}
diamonds %>% 
  sample_n(100) %>% 
plot_ly(data = ., x = ~carat, y = ~price, type = "scatter")
```

```{r}

plot_ly(data = diamonds.sample, x = ~cut, y = ~price) %>% 
  add_trace(p=., type = "scatter")

```

###Styling


```{r}

#specifying scatter, restricting axes and labeling

plot_ly(data = diamonds.sample, x = ~carat, y = ~price) %>% 
  add_trace(p = ., type = "scatter") %>% 
  layout(xaxis = list(range = c(2.5,5), title = "Carat"),
         yaxis = list(title = "Price"))

```

Let's improve upon this by adjusting the text that appears when you hover over a point.

```{r}

#change the text when you hover over points

plot_ly(data = diamonds.sample, x = ~carat, y = ~price, 
        text = ~paste0("Price: ", price, "<br> Carat: ", carat)) %>% 
  add_trace(p = ., type = "scatter") %>% 
  layout(xaxis = list(range = c(2.5,5), title = "Carat"),
         yaxis = list(title = "Price"))

```

```{r}
#create fake data
lines.data <- data.frame(x = c(1:100),
                         trace0 = rnorm(100, mean = 5)) %>% 
  mutate(trace1 = log(trace0),
         trace2 = trace0^2)

plot_ly(data = lines.data, x = ~x) %>% 
  add_trace(y = ~trace0, name = "Trace Zero", mode = "lines") %>% 
  add_trace(y = ~trace1, name = "Trace One", mode = "lines+markers") %>% 
  add_trace(y = ~trace2, name = "Trace Two", mode = "markers")            

```

Trying out some transformations

```{r}

lines.data.wide <- gather(lines.data, "trace", "number", -x)

plot_ly(data = lines.data.wide, x = ~x, y = ~number, color = ~trace, mode = "line")

```

```{r}
plot_ly(data = lines.data.wide, x = ~x, y = ~number, color = ~trace, mode = "line") %>% 
  layout(xaxis = list(ticks = "outside",
                      dtick = 10,
                      ticklen = 5,
                      tickwidth = 2,
                      tickcolor = toRGB("pink"),
                      title = "X",
                      showgrid = FALSE,
                      zeroline = FALSE),
         yaxis = list(ticks = "outside",
                      dtick = 5,
                      ticklen = 5,
                      tickwidth = 1,
                      tickcolor = toRGB("pink"),
                      showgrid = FALSE,
                      zeroline = FALSE,
                      title = "Trace"),
         showlegend = TRUE,
         legend = list(x = .1, y = 5),
         title = "Title")

```

Double click on a legend entry to select just that one.
OR single click on legend to hide it.

```{r}

cnt <- with(diamonds.sample, table(cut, clarity))

plot_ly(data = diamonds.sample, x = ~cut, y = ~clarity, z = ~cnt) %>% 
  add_trace(type = "histogram2dcontour",
            contours = list(showlabels = T, 
                            labelfont = list(color = 'white')))

```

Box Plots!

```{r}

plot_ly(data = diamonds.sample, y = ~price, color = ~cut, type = "box")

#flip orientation
plot_ly(data = diamonds.sample, x = ~price, color = ~cut, type = "box")

```

### 3D Finally

```{r}

plot_ly(data = diamonds.sample, x = ~carat, y = ~price, z = ~depth, color = ~cut)

```

Going back to ggplot2 - build it in ggplot and then convert it to plotly

```{r}
fig1 <- diamonds.sample %>% 
  ggplot(data = ., aes(x = price, y = depth, color = cut))+
  geom_point()+
  stat_smooth()

ggplotly(fig1)
```

Homework 4:
#new dataset, use plotly
